### 2013-12-13 manuscript.tex
added ignore file, added first scheleton of paper

% Created 2013-12-13 Fri 11:53
\documentclass[]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{hyperref}
\tolerance=1000
\author{Daniele Barchiesi}
\date{}
\title{Dictionary learning for classification: learning incoherent subspaces}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs 24.3.1 (Org mode 8.2.4)}}
\begin{document}

\maketitle

\section{Abstract}
\label{sec-1}
\section{Introduction}
\label{sec-2}
\section{Background}
\label{sec-3}
\section{Incoherent Dictionary Learning}
\label{sec-4}
\section{Learning Incoherent Subspaces}
\label{sec-5}
\section{Numerical Experiments}
\label{sec-6}
\section{Discussion}
\label{sec-7}
\section{Conclusions}
\label{sec-8}
% Emacs 24.3.1 (Org mode 8.2.4)
\end{document}